Qualitative Probabilistic Programming

### 2015-06-05 paper/paper.tex
stuff

  The first cluster is at (6.3, 5.0) and the second is at (5.0, 1.6).  They both
  have a standard deviation of 0.54 in the x direction and 0.69 in the y direction.
  We could use these parameters to fill in the generative model:
  \begin{verbatim}
    return repeat(n, function(i) {
      var cluster = randomInteger(nclusters);
      return [gaussian(cluster == 0 ? 6.3 : 5.0, 0.54),
              gaussian(cluster == 0 ? 5.0 : 1.6, 0.69)];
    });
  \end{verbatim}
  This model is estimated to assign probability density $e^{-393}$ to the data, or 0.0728 per point.

  Note that, because the two clusters are forced to have the same variance, they do not fit the data well,
  since the data has a different shape in each location.  To fix this problem, we can substitute the following
  model, which uses a separate random function for each cluster:
  \begin{verbatim}
var Cluster = Categorical(2);
var Point = Vector(2, Double);
var getPointFunctions = [randFunction(Point), randFunction(Point)];
return repeat(n, function(i) {
  var cluster = randomValue(Cluster);
  observe('point' + i, getPointFunctions[cluster]);
});
\end{verbatim}
  Using this model, we get the following clusters:
  \begin{center}
    \includegraphics[scale=0.5]{../plots/irisclusters_indep.png}
  \end{center}
  This model (with the parameters filled in) is estimated to assign probability density $e^{-337}$ to the data, or 0.1058 per point, so it assigns significantly higher likelihood to the data.

@@

  This model is estimated to assign probability $e^{-269}$ to the data, or 0.166 per point, a large improvement from the previous model.

@@

\subsection{A more complex model}

TODO: if there is time, we should put a more complex data science type example, where we add dependencies and show change in accuracy as we add/remove assumptions.

>>>>

  The first cluster is at (6.3, 5.0) and the second is at (5.0, 1.6).  They both
  have a standard deviation of 0.54 in the x direction and 0.69 in the y direction.
  We could use these parameters to fill in the generative model:
  \begin{verbatim}
    return repeat(n, function(i) {
      var cluster = randomInteger(nclusters);
      return [gaussian(cluster == 0 ? 6.3 : 5.0, 0.54),
              gaussian(cluster == 0 ? 5.0 : 1.6, 0.69)];
    });
  \end{verbatim}
  This model is estimated to assign probability density $e^{-393}$ to the data, yielding a perplexity of $e^{-393/150} = 0.0728$.

  Note that, because the two clusters are forced to have the same variance, they do not fit the data well,
  since the data has a different shape in each location.  To fix this problem, we can substitute the following
  model, which uses a separate random function for each cluster:
  \begin{verbatim}
var Cluster = Categorical(2);
var Point = Vector(2, Double);
var getPointFunctions = [randFunction(Point), randFunction(Point)];
return repeat(n, function(i) {
  var cluster = randomValue(Cluster);
  observe('point' + i, getPointFunctions[cluster]);
});
\end{verbatim}
  Using this model, we get the following clusters:
  \begin{center}
    \includegraphics[scale=0.5]{../plots/irisclusters_indep.png}
  \end{center}
  This model (with the parameters filled in) is estimated to assign probability density $e^{-337}$ to the data, yielding a perplexity of 0.1058.  This is a significantly improved fit.

@@

  This model is estimated to assign probability $e^{-269}$ to the data, yielding a perplexity of 0.166.  This is a large improvement from the previous model.

@@
